Towards a better predictive model from rest fMRI: benchmarks across multiple phenotypes

### 2017-06-15 rs_study/experiments/paper/paper.tex
Writing Methods part for paper

%\subsection{Resting State fMRI (rs-fMRI) prediction pipeline}
\autoref{fig:pipeline}  depicts  our  prediction  pipeline consists of
defining brain ROIs, connectivity parameterization and classification using
machine learning. We used this pipeline to study the various choices
at each
step by measuring the impact of each choice made across diverse prediction
tasks from five different rs-fMRI datasets. In below sections, we outline
about the choices used at each step, rs-fMRI datasets with the prediction task
in each dataset and cross-validation framework to prediction accuracy.

%trategy where at step i), we defined brain ROIs from training set of
%rs-fMRI images where the same set is used for training a linear
%model classifier in prediction tasks at step iii).
\subsection{Definition of brain regions of interest (ROIs)}
There exists many different approaches to define brain ROIs as overviewed
in \cite{thirion2014}. Here, we choose to
define brain ROIs using pre-defined brain atlases based on
anatomical and functional and using data-driven methods. For each type of
defining brain ROIs, we chose six variety of various
choices out of which three
pre-defined atlases - Automated Anatomical Labeling (AAL) - Harvard Oxford
- Bootstrap Analysis of Stable Clusters (BASC) and using four data-driven
methods - Linear decomposition models: Group Independent Component
Analysis (Group ICA) - Online Dictionary Learning (DictLearn) and
Clustering models: KMeans - Agglomerative with Ward criterion. For each
data-driven method, we build the atlas on training data ($75\%$) on each
split (n=100) and use this atlas for the consecutive step.

@@

\begin{figure}
    \centerline{%
	\includegraphics[width=0.5\paperwidth]{figures/brainrois.pdf}%
    }%
    \caption{\textbf{Brain regions/parcellations using ICA, DictLearn, KMeans,
        Ward}
    %
    %
    \label{fig:brain_rois}
    }
\end{figure}
\paragraph{Choice of brain atlas}

\section{Discussion:}

@@

\section{Conclusion:}

@@

\section{Supplementary material}
%\bibliographystyle{elsarticle-num-names}
\bibliography{biblio}

>>>>

%\subsection{Resting State fMRI (rs-fMRI) prediction pipeline}
\autoref{fig:pipeline}  depicts  our  functional-connectivity based
prediction  pipeline \citep{abraham2017deriving} and consists of
following steps: i) defining brain ROIs and extraction of timeseries
signals from these ROIs,
ii) connectivity parameterization between pair of ROIs and
iii) connectivity-based classification of phenotypic target using
machine learning.
We study various methods at each
step in the pipeline by measuring the impact of each method
made in predicting diverse phenotypic targets across five different
rs-fMRI datasets.

In below sections, we describe about the methods used at each step
in the pipeline, rs-fMRI datasets and the phenotypic target used
within each dataset and cross-validation framework to reporting
prediction accuracy scores.

As the methods used in the pipeline are well known therefore explaining the
technical details is out of reach in this paper. The main intention here is to
tell how we used the methods rather than what each method is.
%trategy where at step i), we defined brain ROIs from training set of
%rs-fMRI images where the same set is used for training a linear
%model classifier in prediction tasks at step iii).
\subsection{Definition of brain regions of interest (ROIs)}
The first step of the pipeline consists in reducing the
dimensionality of the problem by aggregating whole brain voxels into
each set of voxels forming a brain ROIs. Each ROI has a meaningful structure
underlying in it where we can use this structure to learn the connectivity
interactions to study the classification of phenotypic targets.
There exists many different kinds of approaches to define brain ROIs
as overviewed in \citep{thirion2014}. We choose to define brain ROIs
using two specific approaches. One approach is using reference atlases
where ROIs are pre-defined while the another approach is
using data-driven methods where ROIs are defined by directly learning
on the data in a \textbf{multi-subject level}. For each approach, we used
various choices in reference atlases and data-driven methods to study the
impact of each choice in the prediction pipeline on diverse phenotypic targets.
The question underlying ROIs selection is whether different brain
disorders lead to a consistent choice, and whether genericity
should be preferred to adaptative strategies.

\paragraph{Choices of pre-defined atlases}
We considered three reference atlases: i) \textbf{Automated Anatomical Labeling
(AAL)} \citep{tzourio-mazoyer2002a}, a structural atlas obtained on
unseen single-subject high-resolution T1 image of 116 ROIs,
ii) \textbf{Harvard Oxford}
\citep{desikan2006}, a probabilistic cortical segmented structural
atlas learned on unseen 40 subjects T1 images, contains of 48 ROIs in
each hemisphere of total 96 ROIs. iii) \textbf{Bootstrap Analysis of
Stable Clusters (BASC)} \citep{bellec2010}, a multi scale functional atlas
extracted using K-Means clustering on rs-fMRI dataset which
consists of 43 healthy subjects. We considered exploring various scales set
in $\{036, 064, 122, 197, 325, 444\}$.

\paragraph{Choices of data-driven atlases}
We considered four popular data exploratory methods most widely used in the
fMRI community to learn the functional connectivity of the human brain driven
by intrinsic activity from these ROIs \citep{yeo2011, kahnt2012, thirion2014}
\citep{calhoun2001a, beckmann2004, varoquaux2011, abraham2013}.
Inspired from these research methods, we choose to build brain
atlases from multi-subject rs-fMRI images using \textbf{two clustering
methods} mainly:
i) \textbf{K-Means} \citep{sculley2010}, and ii) hierarchical agglomerative
clustering using \textbf{Ward’s algorithm} \citep{ward1963} added with
spatial connectivity matrix to map together all neighboring voxels sharing
the same information, and \textbf{two
are data decomposition linear methods} where these two methods have quite
different approach in extraction of brain ROIs, namely: iii) \textbf{
Canonical Independent Component Analysis (GroupICA) or (CanICA)}
\citep{varoquaux2010}, iv) \textbf{Online Dictionary Learning
(DictLearn)} \citep{mensch2016}.

\paragraph{Choices of dimensionalities and dimensionality reduction}
Prior to defining brain ROIs with clustering methods
i.e., K-Means and Ward, we reduced the dimensionality
of the multi-subject rs-fMRI images in a group level using Principal Component
Analysis (PCA) \citep{smith2014} to $n=100$ components.
Given these PCA components, we define the brain ROIs with
varying in cluster dimensions in $dim=\{40, 60, 80, 100, 120, 150, 200, 300\}$
where our intention also lies in studying the
impact of dimensionality in building brain ROIs used for prediction of
diverse phenotypic targets.

While with data decomposition methods i.e., CanICA and DictLearn,
we vary the dimensionality reduction components $n$
according to brain ROIs defining dimensions in $dim=\{40, 60, 80, 100, 120,
150\}$\footnote{For less ambiguity, we focussed to report results with less
    dimensions especially for CanICA and DictLearn where the computational
    cost is high in defining brain ROIs. However, we also investigated
for higher dimensions in $dim=\{200, 300\}$ without any improvements in the
results}.
We share almost common parameters across each pair of data-driven methods
to compare them efficiently between clustering methods or
between data decomposition methods. Also, all data-driven methods altogether.

For each data-driven method, we
learn brain ROIs on training set of rs-fMRI images to overcome overfitting
results \citep{abraham2017deriving}.
The ratio of training set used to learn brain ROIs is described below in section
Experimental study. We also used Gaussian smoothing on rs-fMRI images
with Full Width Half Maximum (FWHM) of $6mm$ for all
data-driven methods prior to learning brain ROIs. We used common smoothing to
avoid possible variance with the definition of ROIs which may arise while
varying the smoothness on the rs-fMRI images.

\paragraph{Choices of regions or networks}
While defining brain ROIs, we are also interested to study the
impact on the prediction driven by brain ROIs defined with either regions
or networks. Regions implies separating the
biggest connected components or networks in the brain extracted to independent
regions which are
spatially contiguous \citep{abraham2014region}. For example, separating
bi-hemispheric brain network to regions onto each hemisphere or separating default
mode network (DMN) to each a independent region.
Whereas networks implies using directly as it is without implying any regions
method.

We used a Random Walker segmentation approach to extract regions in the
brain networks decomposed by methods in CanICA and DictLearn as proposed in
\citep{abraham2014region}. While for BASC and K-Means, we used simple
label\footnote{For label reassignment approach, we used multidimensional
module named as ndimage.label from scipy software.}
reassignment approach to extract the biggest connected components to regions
assigned with new label. In particular we compare in regions or networks with
all methods except AAL, Harvard Oxford and Ward.
During this procedure, we remove
spurious regions of size $< 1500 mm^{3}$. An example of the brain atlases obtained
through data-driven methods on ADNI rs-fMRI dataset is presented in
\autoref{fig:brain_rois}.

\paragraph{Time-series signals extraction}
After definition of brain ROIs, we extract a representative time-series
for each ROI in each subject. For atlases composed of non-overlapping
ROIs as can be seen in \autoref{fig:brain_rois} (bottom row), we simply
compute the weighted average of the fMRI
time series signals over all voxels within that specific region.
For fuzzy overlapping ROIs, such as the components extracted
by CanICA and DictLearn as shown in \autoref{fig:brain_rois} (top row),
we use a linear model to unmix the
signal in each voxel as the additive decomposition of signals
over several overlapping ROIs.

Standard pre-processing steps are applied on a signal extraction level
to regress out confounds or nuisance signals. We extracted $10$ CompCor
high-variance confounds on the whole brain using similar method as
proposed in \citep{behzadi2007} and also used $6$ motion confounds pre-extracted
during motion realignment in $6$ directions while pre-processing raw rs-fMRI
images\footnote{\url{https://github.com/neurospin/pypreprocess}}.
We have not used any motion
confounds for ABIDE \citep{dimartino2014autism} and
ACPI\footnote{\url{http://fcon_1000.projects.nitrc.org/indi/ACPI/html/}}
rs-fMRI images as the data is already
pre-processed using standard preprocessed initiative protocols.
More details on pre-processing steps used for raw rs-fMRI
images can be in Section Experimental study.
The signal of each region is also
then normalized, detrended and bandpass-filtered between 0.01
and 0.1 Hz. The primary motivation using pre-processing on a signal extraction
level is to reject non-neural information i.e., information corresponding to
acquisition noise or cardiac activity or physiological random noise and get
cleaned timeseries as much as possible to learn the functional-connectivity
interactions which are crucial for classification step in supervised learning
as proposed in \citep{varoquaux2013connectomes}.

@@

\begin{figure}
    \centerline{%
	\includegraphics[width=1.2\linewidth]{figures/rois2.pdf}%
    }%
    \caption{\textbf{Brain regions/parcellations using ICA, DictLearn, KMeans,
        Ward} Brain atlas with Dictionary Learning is chosen with
        $n=60$ resting state networks separated to regions whereas same with
        ICA but $n=80$ resting state networks. For KMeans and Ward, brain
        parcellations are chosen for $n=120$ dimensions.
    %
    %
    \label{fig:brain_rois}
    }
\end{figure}
\paragraph{Choice of brain atlas}

\section{Discussion}

@@

\section{Conclusion}

@@

\section{References}
\small
%\bibliographystyle{elsarticle-num-names}
\bibliography{biblio}
